Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-30 chatbot_classification/aaai_2021/paper.ver_04.tex
version 4.   modified:   paper.tex
  new file:   paper.ver_04.pdf
  new file:   paper.ver_04.tex

  Task-specific  conversational  chatbots  \cite{wen2016network}  can be applied
  into  many practical products. For example, it can share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another example is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments. Regardless of whether these conversations belong to
  single-turn or multi-turn, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is caple of conducting expected action, text reply or accomplish some
  operations.

  Although  task-specific  conversational chatbot already has a wide application
  in various business and industries, it's still a quite challenging task due to
  its  natural  properties of low-resource. First, customers' utterance within a
  conversation  is  usually  quite  short  and  completed  in a few words. Short
  text \cite{song2014short}  are generally more ambiguous in comparison with long
  texts  such  as  paragraphs  or  documents  since  they  don't  contain enough
  contextual  information,  which poses a great challenge \cite{chen2019deep} for
  classification        task \cite{phan2008learning,yan2009dynamic,hua2015short}.
  Second,  at  the  initial  stage  of  building  a chatbot for specific task or
  service, it's often extremely hard to collect sufficient data samples for each
  class.  The  reason  is  that it's usually too expensive and time consuming to
  extract  different  ways  of natural language expression for each intention class
  from  history  conversation  log  or  even  manually  compose  the corpus from
  nothing.  Therefore,  the  key  to  building a task-specific chatbot with high
  performance     becomes     solving     the     challenge     of    short-text
  classification \cite{sriram2010short}        problem       under       few-shot
  setting \cite{yu2018diverse}.

  One  of  the most popular approach among existing work was text classification
  model  based  on  neural network. Since the length of the text is quite short,
  many   previous   work \cite{wen2016network}  choose  neural  network  such  as
  convolutional                          neural                         networks
  (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very}  or  long
  short  term memory networks (LSTMs) \cite{mousa2017contextual,liu2016recurrent}
  to  accomplish  the task of extracting semantic feature from limited amount of
  words.  A  common model structure is adding a softmax classifier to the top of
  the  neural  network.  Afterwards, pre-trained language models on large corpus
  like   BERT \cite{devlin2018bert}  and  RoBERTa \cite{liu2019roberta}  has  been
  proven   more   powerful  in  solving  many  NLP  tasks  including  short-text
  classification \cite{madabushi2020cost}.      Especially      for      few-shot
  scenarios \cite{yu2018diverse},       pre-trained      model      based      on
  transformers \cite{vaswani2017attention},   shown  in  Fig.  \ref{Fig.1.sub.1},
  tends  to do more help to reducing the negative effects brought by scarcity of
  training data.

@@

  The  third  popular  approach  among  previous work was based on sentence-pair
  model.  The  motivation  of  this  approach  was  started  from  the  idea  of
  Information     Retrival    (IR)    based    chatbot \cite{jafarpour2010filter,
  leuski2011npceditor}.  Having  a  Q-A (Question-Answer) pairs dataset and user
  query  Q,  the  IR based conversational system will look up in the Q-A dataset
  for  the pair (Q', A') that best matches query Q through semantic analysis and
  returns   A'   as   the  answer  to  Q \cite{mnasri2019recent}.  In  this  way,
  sentence-pair  model  pre-trained  on  large  corpus  of  semantic  similarity
  identification  task  can  be applied for being used as a tool to identify the
  class  with  highest  semantic  similarity to customer's query. A common model
  structure  of pre-trained \cite{devlin2018bert} sentence-pair model is based on
  multiple  cross-attention  mechanism \cite{barkan2020scalable},  shown  in Fig.
  \ref{Fig.1.sub.2}.  Due  to  the  fact that many experiment results have shown
  that  RoBERTa \cite{liu2019roberta}  is  an  improved  version of BERT, and has
  achieved  amazing  results  on  both  text  classification  and sentence-pairs
  semantic similarity tasks, we choose RoBERTa as an important baseline approach
  in this paper.

  Despite the success of these 3 approaches, they still have some limitations in
  task-specific chatbot scenario. As for the text classification model approach,
  it's   quite   hard  to  use  data  augmentation  method  to  solve  the  data
  insufficiency  challenge since the it's usually unfeasible to get large amount
  of  domain-specific  data  when  facing  a  new  task. As for label-word joint
  embedding  approach,  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.  With  respect  to  the sentence-pair model approach, though we can
  obtain   large   amount  of  data  in  semantically  duplicate  sentence  pair
  identification  domain  for  transfer  learning \cite{sun2019fine},  it's still
  quite  hard  to  perform  well  on  task-specific dataset because the training
  objective  of  sentence-pair  model  is semantic similarity, which is slightly
  different from the target of classification task. That is to say, there always
  exists  some intention classes which cannot be distinguished by each other merely
  depending on semantic similarity. For example, we have 2 user query saying, A:
  What  should I do if I want to change my password? B: What is the modification
  rule if I want to change my password? In this case, sentence A is semantically
  similar  to  B,  since  they  both  express  the desire for changing password.
  However, it's still reasonable to classify them into 2 intention classes, since A
  is asking for the method for changing password, while B is asking for the rule
  to follow when creating a new password.

  The  above  limitation  motivates  us  to  propose a joint system of both text
  classification  task  and  sentence-pair  semantic  similarity task, named SFC
  here, shown in Fig. \ref{fig:ver.1.3}. Our goal is to utilize the advantage of
  the  feasibility  of  transfer learning based on semantic similarity, while in
  the  mean time, add the classification task's target into the training process
  of            sentence-pair           model           for           multi-task
  learning \cite{caruana1993multitask,collobert2008unified}.  To  obtain  such  a
  joint  system,  we  first start with preparation work, which is pre-training a
  sentence-pair   model   on  external  corpus  for  sentence  pair  duplication
  identification,  and  then  find  an auxiliary model or tool which can help us
  sample out top-K most related intention classes. This auxiliary model or tool can
  either  be searching engine such as elasticsearch \cite{divya2013elasticsearch}
  or  a  text  classification  model  trained  on  task-specific  chatbot  data.
  Afterwards,  we  further  fine-tune  the  sentence-pair  model with 2 training
  tasks:  1)  the  regular  sentence pair similarity tasks. Here we use the text
  classification  model  as  an auxiliary model in this paper to sample sentence
  pairs       for       training       based      on      negative      sampling
  strategy \cite{bamler2020extreme}.  The  reason is that a task-specific chatbot
  often  has hundreds of intention classes, which forms too many sentence pairs for
  training  since  any  two  sentences  from  two  different  classes can form a
  sentence-pair     of     negative     sample.     Besides,     according    to
  Bamler \cite{bamler2020extreme},   training   on  negative  samples  from  most
  confusing  wrong  label  can  help  model  converge  faster  and obtain better
  performance.  2) the classification task on top-K (here K is a hyperparameter)
  candidate  classes  provided  by  our auxiliary model or tool. Here we use the
  average  pooling  of  the  representation  given  by  sentence-pair  model for
  sentence pairs formed by the user input query and the corpus sentences in each
  class  as  the  feature.  We  stack the task-specific layers on the top of the
  shared-parameter  sentence-pair  model structure, which is a classic parameter
  sharing                                   mechanism \cite{collobert2008unified,
  subramanian2018learning,liu2019multi}   for  multi-task \cite{sun2019fine}.  In
  this  way, specific knowledge contained in these 2 tasks can be fully explored
  and deeply interact with each other to obtain better overall performance.

  However,  since  the auxiliary tool in joint system mentioned above works in a
  separate  stage  from  sentence-pair model, we find the the quality of sampled
  candidate  classes  might  limit  the  overall  performance  of SFC, since the
  candidate  classes  for  each user query is always fixed during the multi-task
  training  process. This observation motivates us to further improve SFC into a
  joint  model  of  sentence-pair  model  and  text  classification  model  in a
  joint  architecture  by  putting  different  tasks at different network
  layers \cite{sogaard2016deep,   hashimoto2016joint}   and  then  training  them
  together,  shown  in  Fig  \ref{fig:framework}.  Here, the text classification
  model  involved in joint training process replace the role of auxiliary tools.
  In  this  way, the sentence pairs selected from top-K classes becomes dynamic,
  which  means  the  text classification model will also be further optimized to
  provide better top-K classes pooling result.

  We summarize our contributions as follow:

  1)  We  propose  a  novel  joint  system  named  SFC  with multi-task training
  technique designed for task-specific conversational chatbot with low-resource,
  in  which  we make full utilization of the advantages brought by both the text
  classification task and sentence-pair semantic similarity task.

  2)  To  further  improve  the  system  performance  of SFC, we also propose an
  innovative  joint  model structure, in which the text classification model and
  sentence-pair model are fused into one single model for joint training.

  3)  Experiment  results  on  4  public and 1 private short-text classification
  datasets with respect to intention recognition tasks under conversational chatbot
  scenario  all demonstrated that our proposed SFC joint system with multi-task,
  as  well  as  the  improved  version  of  SFC,  can  both  achieve  remarkable
  improvement comparing to some powerful single-task and single-model baselines,
  especially under few-shot settings.

@@

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 14.25 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intention detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intention, joint SFC still outperforms LEAM by
  14.69  percents  on  average  F-score.  This  observation  indicates that it's
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.

@@

  \textbf{Settings  of  hyperparameters.} Finally, we analyze the effect brought
  by  different  settings  of  hyperparameters,  see  Table \ref{tbe:table3}. We
  conduct  experiments on 5 different settings of hyperparameters, which are the
  number  of candidate class number $K$ and the number of sentence pairs $P$ for
  each  class, on ITG and Amazon-670K datasets to find the most suitable setting
  of  these  2  hyperparameters.  We  control the total number of sentence pairs
  sampled  for  each  user  query  into  about  the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be slightly different for
  datasets of different data size. The class label number of ITG and Amazon-670k
  are  similar  in  our  experimental  setting, which is around 250. However the
  total  amount  of  data  points  of  ITG  is approximately 2 times larger than
  Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG
  since  most  of  the  classes  contains  over  10  data  points.  However, for
  Amazon-670k,  the amount of data sample for each class is scarce, thus, we may
  need  more  candidate  classes  to  sample enough effective sentence pairs for
  sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$
  becomes a good choice. However, in spite of the effect on evaluation performance brought by different settings of hyperparameters, joint SFC still steadily outperforms two-stage SFC by 0.41 percents in overall average F-score on ITG and Amazon-670k datasets.

>>>>

  Task-specific  conversational  chatbot  \cite{wen2016network}  has been applied
  into  many practical products. For example, it can share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another example is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments. Regardless of whether these conversations belong to
  single-turn or multi-turn, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is caple of conducting expected actions, text reply or accomplish some
  operations.

  It  is  a  challenging task due to some inherent properties. First,
  a customer's  utterance  within  a conversation is usually quite short. Since it
  does  not  contain  enough  information,  short  text \cite{song2014short} are
  thought  to  be  more  ambiguous  compared to long texts such as paragraphs or
  documents,  poseing  a  great challenge \cite{chen2019deep} for classification
  task   \cite{phan2008learning,yan2009dynamic,hua2015short}.   Second,  in  the
  initial  stage  of  building  a  chatbot  , it is often rather hard to collect
  sufficient  labeled  data  to  train  a good model. Because developers have to
  examine tons of real system logs for typical user's utterances, and then label
  them  with expected intentions. Therefore, the key to building a task-specific
  chatbot  with  high  performance  becomes  solving the challenge of short-text
  classification   \cite{sriram2010short}   problem   under   few-shot   setting
  \cite{yu2018diverse}.

  Generally  there  are  two  kinds  of  approaches  to  address  this  problem,
  classification model that directly maps an input into a most likely label, and
  similarity  model  that searches a label whose associated data is most similar to
  an input.

  The   first  one,  \emph{text  classification},  includes  classic  long  text
  classification  and  short  text  classification  in  this  scenario.  Besides
  traditional  machine  learning  models  like  SVM,  boosting  tree,  many work
  \cite{wen2016network}  choose  neural  network  such  as  convolutional neural
  networks (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very}
  and       long       short       term       memory       networks      (LSTMs)
  \cite{mousa2017contextual,liu2016recurrent}   to   accomplish   the   task  of
  extracting  semantic  feature from limited amount of words. Their common model
  structure  is  adding  a  softmax classifier as the final layer and mapping to
  labels.  Another  two  interesting lines of work, label-word joint models, and
  joint  NER  and  classification  are  discussed  in  the related work section.
  Afterwards,   pre-trained   language   models   on   large  corpus  like  BERT
  \cite{devlin2018bert}  and  RoBERTa \cite{liu2019roberta} has been proven more
  powerful  in  solving  many  NLP  tasks  including  short-text  classification
  \cite{madabushi2020cost}.      Especially      for      few-shot     scenarios
  \cite{yu2018diverse},     pre-trained     model    based    on    transformers
  \cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends to do more
  help to reducing the negative effects brought by scarcity of training data.

  The  second  one,  \emph{similarity model} is also called sentence-pair model.
  Its  motivation  originates  from  the idea of Information Retrival (IR) based
  chatbot   \cite{jafarpour2010filter,   leuski2011npceditor}.   Having   a  Q-A
  (Question-Answer)  pairs dataset and user query Q, the IR based conversational
  system will look up in the Q-A dataset for the pair (Q', A') that best matches
  query  Q  through  semantic  analysis  and  returns  A'  as  the  answer  to Q
  \cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large
  corpus  of  semantic  similarity  identification task can be applied for being
  used  as  a  tool  to  identify  the class with highest semantic similarity to
  customer's    query.    A    common    model    structure    of    pre-trained
  \cite{devlin2018bert} sentence-pair model is based on multiple cross-attention
  mechanism  \cite{barkan2020scalable},  shown in Fig. \ref{Fig.1.sub.2}. Due to
  the   fact   that   many   experiment   results   have   shown   that  RoBERTa
  \cite{liu2019roberta} is an improved version of BERT, and has achieved amazing
  results  on  both  text  classification and sentence-pairs semantic similarity
  tasks, we choose RoBERTa as an important baseline approach in this paper.

  Despite  the  success  of  these two kinds of approaches, they still have some
  limitations   in   task-specific  chatbot  scenario.  As  for  the  \emph{text
  classification  model},  it  is  quite hard to use data augmentation method to
  solve  the  data insufficiency challenge since the it is usually unfeasible to
  get  large amount of domain-specific data when facing a new task. With respect
  to  the  \emph{similarity model}, though we can obtain a large amount of data
  from semantically  duplicate  sentence  pair identification domain for transfer
  learning  \cite{sun2019fine},  it  is  still  quite  hard  to  perform well on
  a new task-specific dataset,  since their objective losses are totally different.

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:ver.1.3}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elasticsearch  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

  We summarize our contributions as follow:
  \begin{enumerate}
    \item We propose a novel joint system to makes full utilization of
    the advantages from both text classification model and similarity model.

    \item To  better  train such a system, we bring in multi-task training to obtain
    expected performance.

    \item Experiment  results  from 4 public and 1 private short-text classification
    datasets,  show  that  our  proposed  SFC joint system can achieve significant
    improvements over strong baselines, especially in the low resource settings.
  \end{enumerate}

  \section{Related work}

@@

  As for \emph{label-word joint
  embedding  approach},  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.

  {Add joint NER and classification problem here.}

@@

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 14.25 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intention detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intention, joint SFC still outperforms LEAM by
  14.69  percents  on  average  F-score.  This  observation  indicates that it is
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.

@@

  \textbf{Settings  of  hyperparameters.} Finally, we analyze the effect brought
  by  different  settings  of  hyperparameters,  see  Table \ref{tbe:table3}. We
  conduct  experiments on 5 different settings of hyperparameters, which are the
  number  of candidate class number $K$ and the number of sentence pairs $P$ for
  each  class, on ITG and Amazon-670K datasets to find the most suitable setting
  of  these  2  hyperparameters.  We  control the total number of sentence pairs
  sampled  for  each  user  query  into  about  the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for
  datasets of different data size. The class label number of ITG and Amazon-670k
  are  similar  in  our  experimental  setting, which is around 250. However the
  total  amount  of  data  points  of  ITG  is approximately 2 times larger than
  Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG
  since  most  of  the  classes  contains  over  10  data  points.  However, for
  Amazon-670k,  the amount of data sample for each class is scarce, thus, we may
  need  more  candidate  classes  to  sample enough effective sentence pairs for
  sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$
  becomes a good choice. However, in spite of the effect on evaluation performance brought by different settings of hyperparameters, joint SFC still steadily outperforms two-stage SFC by 0.41 percents in overall average F-score on ITG and Amazon-670k datasets.
